\section{\label{sec:intro}Introduction}

The Hamiltonian of mean force (HMF) is the operator whose Gibbs form reproduces the reduced equilibrium state of a quantum system coupled to an environment~\cite{campisiFluctuationTheoremArbitrary2009,talknerColloquiumStatisticalMechanics2020,trushechkinOpenQuantumSystem2022,seifertFirstSecondLaw2016}. For a total Hamiltonian $H = H_S + H_B + f\otimes B$, the unnormalised reduced equilibrium operator is $\bar\rho_S(\beta) = \Tr_B\,e^{-\beta H}$, and the HMF is defined by $e^{-\beta H_{\mathrm{MF}}(\beta)} = \bar\rho_S(\beta)/Z_B(\beta)$. At weak coupling, $H_{\mathrm{MF}}$ reduces to the bare system Hamiltonian; at finite coupling it acquires temperature-dependent operator content that is not captured by a simple renormalisation~\cite{hanggiFiniteQuantumDissipation2008,ingoldSpecificHeatAnomalies2009,espositoNatureHeatStrongly2015,rivasStrongCouplingThermodynamics2020}.

The influence-functional formalism provides a natural route to the HMF~\cite{feynmanTheoryGeneralQuantum1963a,caldeiraQuantumTunnellingDissipative1983a,grabertQuantumBrownianMotion1988}. For Gaussian baths with linear coupling, the bath degrees of freedom can be integrated out exactly, producing a bilocal imaginary-time kernel whose Hubbard--Stratonovich (HS) decoupling~\cite{hubbardCalculationPartitionFunctions1959a,stratonovich1957QDistro} represents the reduced equilibrium state as a Gaussian average over quenched propagators~\cite{mccaulPartitionfreeApproachOpen2017c,mccaulDrivingSpinbosonModels2018a,stockburgerExactNumberRepresentation2002}. In the commuting sector $[H_S,f]=0$, the time-ordering obstruction vanishes and the Gaussian average can be performed in closed form, yielding an exact HMF~\cite{campisiTalknerHanggi2009Solvable}. This programme was carried out explicitly in Ref.~\cite{mccaulPartitionfreeApproachOpen2017c} for harmonic baths, where the resulting HMF contains a single bath-induced correction $-\lambda f^2$ with $\lambda$ the reorganisation energy.

However, many physically important environments are not Gaussian. Spin baths~\cite{ProkofevStamp2000}, collections of two-level fluctuators, and any bounded bath generate non-Gaussian statistics in the collective coupling variable~\cite{HsiehCao2018I,HsiehCao2018II}. When the individual couplings remain $\mathcal{O}(1)$ as the bath grows, the central limit theorem does not apply and connected correlators of order $n\geq 3$ contribute at leading order. The influence functional then contains higher-order multilocal terms beyond the Gaussian kernel, and the single HS transformation is insufficient.

Several approaches address non-Gaussian effects. Cluster and cumulant expansions have been applied to spin-bath dynamics~\cite{SuarezSilbey1991,Makri1999,YingEtAl2024}, and the polaron and reaction-coordinate frameworks absorb part of the interaction into a renormalised system~\cite{leggettDynamicsDissipativeTwostate1987,HsiehCao2018II}. Functional-integral treatments of non-Gaussian noise have appeared in real-time settings~\cite{FunoIshizaki2024,HalataeiEtAl2025}, and recent work has examined the structure of the HMF beyond harmonic baths~\cite{burkeStructureHamiltonianMean2024,duGeneralizedHamiltonianMeanforce2025a}. What is missing is a constructive, exact representation of the full non-Gaussian influence functional that reduces to a tractable operator expression for the HMF.

In this paper we provide such a construction. We show that the complete imaginary-time cumulant expansion of the influence functional---including all even-order connected correlators---admits an exact representation as a nested hierarchy of Gaussian auxiliary fields. At order $2n$, the $n$-local composite of system coupling operators is paired with a Gaussian field whose covariance is determined by the symmetry-projected cumulant kernel $K_{2n}^{(n|n)}$. The construction is not perturbative: it is an exact identity at each order, valid for kernels that are indefinite, non-positive-semi-definite (non-PSD), or complex. Contour rotation of the auxiliary fields mode-by-mode suffices to define the Gaussian integrals.

In the commuting sector, the entire nested hierarchy re-averages analytically. Each auxiliary field produces a scalar Gaussian contraction, and the resulting series exponentiates to give an exact closed-form HMF:
\begin{equation}
    H_{\mathrm{MF}}(\beta) = H_S - \sum_{n\geq 1} \alpha_{2n}(\beta)\,f^{2n},
    \label{eq:HMF_preview}
\end{equation}
with $\alpha_{2n} = \frac{1}{(2n)!}\int_0^\beta d\tau_1\cdots d\tau_{2n}\,K_{2n}(\tau_1,\ldots,\tau_{2n})$. This is the main result of the paper.

The paper is structured as follows. Section~\ref{sec:cumulants} defines the reduced equilibrium operator and derives the ordered cumulant expansion of the influence functional. Section~\ref{sec:nested_hs} constructs the nested HS representation for arbitrary even-order cumulants and states the general theorem. Section~\ref{sec:commuting} proves the commuting-sector theorem. Section~\ref{sec:numerics} validates the construction via discretised functional sampling for a qutrit clock model and a multi-qubit magnetisation model. Section~\ref{sec:consequences} analyses the consequences of the polynomial HMF structure for qubits, qutrits, and integrable spin chains.
