\section{\label{sec:nested_hs}Nested Hubbard--Stratonovich Construction}

The Gaussian ($K_2$) sector is decoupled by a single HS field because it is quadratic in the coupling operator. The higher-order analogue is to rewrite each even cumulant contribution as a quadratic form in a composite field, then apply HS at that level.

\subsection{Quadratic-form rewriting at order 2n}

Define the $n$-local composite
\begin{equation}
\Psi_n(\tau_1,\dots,\tau_n)=\prod_{j=1}^{n} f(\tau_j).
\label{eq:v2_psin_def}
\end{equation}
The $2n$-th cumulant contribution can be written as
\begin{equation}
\Phi^{(2n)}
=
\frac{1}{2}\,\langle \Psi_n,\,C_n\,\Psi_n\rangle,
\label{eq:v2_phi2n_quadratic}
\end{equation}
where $C_n$ is the symmetry-projected kernel built from $K_{2n}$ and the inner product integrates all time arguments. This is the precise higher-order completion-of-square analogue.

\subsection{Level-wise HS decoupling}

For any symmetric kernel $C_n$ (not necessarily PSD),
\begin{equation}
e^{\frac{1}{2}\langle \Psi_n,C_n\Psi_n\rangle}
=
\mathbb E_{\eta_n}\!\left[
e^{\langle \eta_n,\Psi_n\rangle}
\right],
\label{eq:v2_hs_level}
\end{equation}
with Gaussian auxiliary field $\eta_n$ of covariance $C_n$. Therefore
\begin{equation}
\left\langle \mathcal T_\tau e^{\int Bf}\right\rangle_B
=
\prod_{n\ge 1}
\mathbb E_{\eta_n}
\left[
\exp\!\left(
\frac{1}{n!}\int \eta_n\,\Psi_n
\right)
\right].
\label{eq:v2_nested_hs}
\end{equation}
Equation~\eqref{eq:v2_nested_hs} is exact order by order in cumulants.

\subsection{Indefinite kernels and contour rotation}

If $C_n$ has negative eigenvalues, one rotates those auxiliary components by $\eta\to i\eta$. The Gaussian integral remains convergent mode-by-mode and Eq.~\eqref{eq:v2_hs_level} remains exact. In practice we monitor imaginary-leakage diagnostics
\(
\max |\im w|/(|\re w|+\epsilon)
\)
for sampled weights.

\subsection{Practical truncation used in numerics}

The simulations in this paper use commuting-sector truncation ladders:
\begin{align}
H_{\mathrm{MF}}^{(2)} &= H_S-\alpha_2 f^2, \nonumber\\
H_{\mathrm{MF}}^{(4)} &= H_S-\alpha_2 f^2-\alpha_4 f^4, \\
H_{\mathrm{MF}}^{(6)} &= H_S-\alpha_2 f^2-\alpha_4 f^4-\alpha_6 f^6.\nonumber
\end{align}
These are compared directly against exact reduced states from finite-bath exact diagonalisation.
